Vowel Elision in Yoruba Collocations

### 2023-03-03 paper.tex
Add source, fix formatting

@@

Our analysis so far accounts for the phenomena in \pref{fig:deletion-simple},
but not the phenomena in \pref{fig:deletion-except-1} or
\pref{fig:deletion-except-2}. We proceed, now, to analyze the first of these, in
which [i] is deleted regardless of whether it the morpheme initial or final
sound. One possible explanation as to why this is the case might be some
quality of the vowels themselves--- the frontness or backness of the vowels
appears to have no immediately discernible pattern with respect to deletion, so
we might try vowel height. Indeed, height appears to be correlated: [u] and
[i], which are both high vowels, are involved in the exceptional deletion
cases! We analyze both of these instances through the mechanism of sonority,
which offers an explanation for high vowels being involved in deletion--- they
are less sonorous. The sonority scale of vowels has been discussed in prior
work[\textbf{cite the quality-sensitive stress paper}, wikipedia article?].
This scale is shown in \pref{table:vowel-sonority}.

@@

Heinz, Kobele, and Riggle \cite{heinz2009evaluating} contest this result,
showing that a different representation of \textsc{OT} grammars as a single
finite state transducer (instead of a sequence of finite state constriants)
averts the computational intractability issue demonstrated in
\cite{idsardi2006simple}. Additionally, they outline two subclasses of the
\textit{universal problem} of attempting to determine a valid constraint
ranking and underlying representations given only surface forms. Namely,
the \textit{simple problem} asks for optimal surface forms given an underlying
form and fixed constraint set \textsc{Con} and ranking \textsc{R}. The
\textsc{quasi-universal problem} asks for an optimal surface form given
a fixed constraint set with any possible ranking.

>>>>

\usepackage{setspace}
% \doublespacing

@@

Our analysis so far accounts for the phenomena in \pref{fig:deletion-simple},
but not the phenomena in \pref{fig:deletion-except-1} or
\pref{fig:deletion-except-2}. We proceed, now, to analyze the first of these, in
which [i] is deleted regardless of whether it the morpheme initial or final
sound. One possible explanation as to why this is the case might be some
quality of the vowels themselves--- the frontness or backness of the vowels
appears to have no immediately discernible pattern with respect to deletion, so
we might try vowel height. Indeed, height appears to be correlated: [u] and
[i], which are both high vowels, are involved in the exceptional deletion
cases! We analyze both of these instances through the mechanism of sonority,
which offers an explanation for high vowels being involved in deletion--- they
are less sonorous. The sonority scale of vowels has been discussed in prior
work \cite{kenstowicz1997quality}. This scale is shown in
\pref{table:vowel-sonority}.

@@

Heinz, Kobele, and Riggle \cite{heinz2009evaluating} contest this result,
showing that a different representation of \textsc{OT} grammars as a single
finite state transducer (instead of a sequence of finite state constriants)
averts the computational intractability issue demonstrated in
\cite{idsardi2006simple}. Additionally, they outline two subclasses of the
\textit{universal problem} of attempting to determine a valid constraint
ranking and underlying representations given only surface forms. Namely,
the \textit{simple problem} asks for optimal surface forms given an underlying
form and fixed constraint set \textsc{Con} and ranking \textsc{R}. The
\textit{quasi-universal problem} asks for an optimal surface form given
a fixed constraint set with any possible ranking.
